### 2018-03-17 paper/3-abstractions.tex
More work on S5
Work on S3

\vspace{-2mm}
\paragraph{Build system} A \emph{build system} takes a task description, a target
key, and a store, and returns a new store in which the target key has an
up to date value.

@@

\noindent
The keys that belong to the transitive closure of \hs{key}'s dependencies fall
into two classes: input keys and non-input keys, which we will denote by $I$ and
$O$, respectively. Note that \hs{key} may be in either of these sets, although
the case when \hs{key} is an input is degenerate: we have $I=\{$\hs{key}$\}$ and
$O=\emptyset$.

@@

@@

So much for applicative tasks. What about monadic tasks with dynamic
dependencies? As we have seen in~\S\ref{sec-background-shake}, dynamic
dependencies need to be tracked too. This cannot be done statically; notice that
the application of the function \hs{dependencies} to a \hs{Task}~\hs{Monad} will
not typecheck. We need to run a monadic task on a store with concrete values,
which will determine the discovered dependencies. Accordingly, we introduce
the function \hs{track}: a combination of \hs{compute} and \hs{dependencies}
that computes both the resulting value and the list of its dependencies:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
import Control.Monad.Writer
\end{minted}
\vspace{0.5mm}
\begin{minted}[xleftmargin=10pt]{haskell}
track :: Task Monad k v -> (k -> v) -> k -> Maybe (v, [k])
track task fetch = fmap runWriter . task (\k -> writer (fetch k, [k]))
\end{minted}
\vspace{1mm}

\noindent
\todo{AM}{Type of track has changed - we no longer supply the Store, just an access fuction}
\todo{AM}{trackM is essential for Shake, what do we do there?}
The implementation uses the standard Haskell \hs{Writer} monad, which allows us
to record addition information -- a list of keys of type \hs{[@@k]} -- when
computing a task. We instantiate \hs{f} to \hs{Writer}~\hs{[@@k]}~\hs{v}, and feed
\hs{fetch}~\hs{k}~\hs{=}~\hs{writer}~\hs{(@@getValue}~\hs{store}~\hs{k,}~\hs{[@@k])}
to the task, which in addition to fetching a value from the store, tracks the
corresponding key. Here is an example of \hs{track}ing monadic tasks:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
@\ghci@ store1 = initialise () (\key -> if key == "A1" then 10 else 1)
@\ghci@ track sprsh2 store1 "B1"
Just (1,["C1","B2"])
\end{minted}
\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
@\ghci@ store2 = initialise () (\key -> if key == "A1" then 10 else 20)
@\ghci@ track sprsh2 store2 "B2"
Just (20,["C1","B1"])
\end{minted}
\vspace{1mm}

\noindent
As expected, the dependencies of \hs{sprsh2} (see the spreadsheet
in~\S\ref{sec-task-monad}) are determined by the contents of the store.
Furthermore, the above example highlights the static cycle between cells
\cmd{B1} and \cmd{B2}: in \hs{store1}, we have \cmd{B1 = B2 = A1 = 10},
whereas in \hs{store2}, we have \cmd{B2 = B1 = A2 = 20}.

>>>>

\vspace{-2mm}
\paragraph{Build system} A \emph{build system} takes a task description, a
\emph{target} key, and a store, and returns a new store in which the target key
and all its dependencies have an up to date value.

@@

\noindent
The keys that are reachable from the target \hs{key} via dependencies fall
into two classes: input keys and non-input keys, which we will denote by $I$ and
$O$, respectively. Note that \hs{key} may be in either of these sets, although
the case when \hs{key} is an input is degenerate: we have $I=\{$\hs{key}$\}$ and
$O=\emptyset$.

@@

The above definition of correctness needs to be adjusted to accommodate
non-deterministic tasks and shallow cloud builds; we address this
in sections~\S\ref{sec-non-determinism} and~\S\ref{sec-cloud-aspects}, respectively.

@@

So much for applicative tasks. What about monadic tasks with dynamic
dependencies? As we have seen in~\S\ref{sec-background-shake}, dynamic
dependencies need to be tracked too. This cannot be done statically; notice that
the application of the function \hs{dependencies} to a \hs{Task}~\hs{Monad} will
not typecheck. We need to run a monadic task on a store with concrete values,
which will determine the discovered dependencies. Accordingly, we introduce
the function \hs{track}: a combination of \hs{compute} and \hs{dependencies}
that computes both the resulting value and the list of its dependencies in an
arbitrary monadic context~\hs{m}:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
import Control.Monad.Writer
\end{minted}
\vspace{0.5mm}
\begin{minted}[xleftmargin=10pt]{haskell}
track :: Monad m => Task Monad k v -> (k -> m v) -> k -> Maybe (m (v, [k]))
track task fetch = fmap runWriterT . task trackingFetch
  where
    trackingFetch :: k -> WriterT [k] m v
    trackingFetch k = tell [k] >> lift (fetch k)
\end{minted}
\vspace{1mm}

\noindent
The implementation uses the standard Haskell \hs{WriterT} \emph{monad
transformer}~\cite{liang1995monad}, which allows us
to record addition information -- a list of keys of type \hs{[@@k]} -- when
computing a task in an arbitrary monad~\hs{m}. We substitute the given \hs{fetch}
with a \hs{trackingFetch} that, in addition to fetching a value, tracks the
corresponding key. The \hs{task} returns the value of type
\hs{Maybe}~\hs{(@@WriterT}~\hs{[@@k]}~\hs{m}~\hs{v)}, which we unwrap by applying
\hs{runWriterT} to the contents of \hs{Maybe}.

% Here is an example of \hs{track}ing monadic tasks:

% \vspace{1mm}
% \begin{minted}[xleftmargin=10pt]{haskell}
% @\ghci@ store1 = initialise () (\key -> if key == "A1" then 10 else 1)
% @\ghci@ track sprsh2 store1 "B1"
% Just (1,["C1","B2"])
% \end{minted}
% \vspace{1mm}
% \begin{minted}[xleftmargin=10pt]{haskell}
% @\ghci@ store2 = initialise () (\key -> if key == "A1" then 10 else 20)
% @\ghci@ track sprsh2 store2 "B2"
% Just (20,["C1","B1"])
% \end{minted}
% \vspace{1mm}

% \noindent
% As expected, the dependencies of \hs{sprsh2} (see the spreadsheet
% in~\S\ref{sec-task-monad}) are determined by the contents of the store.
% Furthermore, the above example highlights the static cycle between cells
% \cmd{B1} and \cmd{B2}: in \hs{store1}, we have \cmd{B1 = B2 = A1 = 10},
% whereas in \hs{store2}, we have \cmd{B2 = B1 = A2 = 20}.
